---
title: "class09"
author: "Phoebe LI"
date: "2/13/2022"
output: pdf_document
---
```{r}
# import the csv file
db <- read.csv("Data Export Summary.csv", row.names = 1)
db

# Xray structure percent
xray.percent <- sum(db$X.ray)/sum(db$Total)*100
round(xray.percent, 2)

# EM structure percent
EM.percent <- sum(db$EM)/sum(db$Total)*100
round(EM.percent, 2)
```


>Q1: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

x-ray:87.2%  EM:5.39%


>Q2: What proportion of structures in the PDB are protein?

```{r}

percent.protein <- (db$Total[1])/sum(db$Total)
round(percent.protein*100, 2)

```


>Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

4486 Structures


>Q4: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?

Because It does not show the H atom in this structure. The resolution is too low that they can not detect H atom.

> Q5: There is a conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have (see note below)?

OH308O


####  Introduction to Bio3D in R
```{r}
install.packages("bio3d", repos="http://cran.us.r-project.org")
library(bio3d)
```

## Reading PDB file data into R

```{r}
pdb <- read.pdb("1hsg")
pdb
```

```{r}
attributes(pdb)
head(pdb$atom)
```

> Q7: How many amino acid residues are there in this pdb object? 

PQITLWQRPLVTIKIGGQLKEALLDTGADDTVLEEMSLPGRWKPKMIGGIGGFIKVRQYD
      QILIEICGHKAIGTVLVGPTPVNIIGRNLLTQIGCTLNFPQITLWQRPLVTIKIGGQLKE
      ALLDTGADDTVLEEMSLPGRWKPKMIGGIGGFIKVRQYDQILIEICGHKAIGTVLVGPTP
      VNIIGRNLLTQIGCTLNF
      
      198 amino acid residues
      
> Q8: Name one of the two non-protein residues? 

 HOH
 
> Q9: How many protein chains are in this structure? 

2 chains





>Q10. Which of the packages above is found only on BioConductor and not CRAN? 

 msa

>Q11. Which of the above packages is not found on BioConductor or CRAN?: 

bio3d-view

>Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket? 

True

>Q13

```{r}
options(repos = list(CRAN="http://cran.rstudio.com/"))
install.packages("bio3d")
install.packages("ggplot2")
install.packages("ggrepel")
install.packages("devtools")
install.packages("BiocManager")

BiocManager::install("msa")
devtools::install_bitbucket("Grantlab/bio3d-view")
```
```{r}
library(bio3d)
aa <- get.seq("1ake_A")
```

```{r}
aa
```

> Q13. How many amino acids are in this sequence, i.e. how long is this sequence? 
214

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
pdbs <- pdbaln(files, fit = TRUE)
```

```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```
### Viewing our superposed structures

```{r}
install.packages("rgl")
library(bio3d.view)
library(rgl)

view.pdbs(pdbs)
```

```{r}
setwd(dudu/BGGN213/class09/PhoebeLi)

```

