---
title: "Class08new"
author: "Phoebe LI"
date: "2/13/2022"
output: pdf_document
---

### Unsupervised learning mini-project

## 1. Let us exploratory data analysis first
```{r}
library(dplyr)
read.csv("WisconsinCancer.csv")
```


```{r}
fna.data <- "WisconsinCancer.csv"
wisc.df <-read.csv(fna.data, row.names=1)
View(wisc.df)

```

```{r}
wisc.data <- wisc.df[,-1]
```
Q1. How many observations are in this dataset?
```{r}
# Have a brief view first
View(wisc.df)
# How many row this data has?
nrow(wisc.df)
```

```{r}
# How many people got diagnosis?
table(wisc.df$diagnosis)
# How many of them are malignant :(
sum(wisc.df$diagnosis=="M")
```

Q3. How many variables/features in the data are suffixed with _mean? 
```{r}
length(grep("mean", colnames(wisc.df)))
# OR
grepl("_mean", colnames(wisc.df)) %>% sum(.)
```


## 2. Principal Component Analysis (PCA)
the main function in base R for PCA is 'prcomp()'. There is an important optional argument called 'scale'
```{r}
diagnosis <- as.factor(wisc.df$diagnosis)
wisc.data<- wisc.df[,-1]
```

```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
```

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale =TRUE)
# Look at summary of results
summary(wisc.pr) 

```


##Interpreting PCA results
```{r}
#plot biplot
biplot(wisc.pr)
```
Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?
It is very difficult to understand.Too much information in a very small scale of plot. it is very overwhelming to read.

```{r}
# Generate plot for principal components 1 and 2
plot(wisc.pr$x[, 1:2], col = diagnosis , 
     xlab = "PC1", ylab = "PC2")
```

Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?
There is less clear boundary between two different colors. 
```{r}
plot(wisc.pr$x[, c(1,3)], col = diagnosis, 
     xlab = "PC1", ylab = "PC3")
```

```{r}
# Create a data.frame for ggplot
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis

# Load the ggplot2 package
library(ggplot2)

# Make a scatter plot colored by diagnosis
ggplot(df) + 
  aes(PC1, PC2, col=diagnosis) + 
  geom_point() + labs(x="PC1", y="PC2")
```
## Variance explained

```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)
```
```{r}
# Variance explained by each principal component: pve
pve <- pr.var/sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

```{r}
## ggplot based graph
install.packages("factoextra", repos="http://cran.us.r-project.org")
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```
Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
44.3%
Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
PC1+PC2+PC3= 72.7%
Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
PC1+PC2+PC3+PC4+PC5+PC7=91.1%

Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?

```{r}
wisc.pr$rotation[,1]
```

Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?
P1-P5, 5 of them

## 3. Hierarchical clustering
Scale the wisc.data data using the "scale()" function

```{r}
#hierachical clustering
data.scaled <- scale(wisc.data)

# Calculate the (Euclidean) distances 
data.dist <- dist(data.scaled)
```


```{r}
wisc.hclust <- hclust(data.dist, "complete")
```

Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

```{r}
# Create a hierarchical clustering model using complete linkage
plot(wisc.hclust)
abline(h=18, col="red", lty=2)
```
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
plot(wisc.hclust.clusters)
table(wisc.hclust.clusters, diagnosis)
```

Q12. Can you find a better cluster vs diagnoses match by cutting into a different number of clusters between 2 and 10?
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=6)
plot(wisc.hclust.clusters)
```
Q13. Which method gives your favorite results for the same data.dist dataset? Explain your reasoning.
```{r}
wisc.pr.hclust <- hclust(data.dist, method="ward.D2")
plot(wisc.pr.hclust)
```
## 5. Combining methods
Clustering on PCA results
```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
```
Q15. How well does the newly created model with four clusters separate out the two diagnoses?

```{r}
wisc.hclust.clusters <- cutree(hclust(data.dist),k = 6)
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)
# Compare to actual diagnoses
table(wisc.pr.hclust.clusters, diagnosis)
```

Q16. How well do the k-means and hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses.
```{r}
table( wisc.hclust.clusters, diagnosis)
```

