---
title: "Class08"
author: "Phoebe LI"
date: "2/11/2022"
output: html_document
---
##Unsupervised learning mini-project

1. Let us exploratory data analysis first
```{r}
library(dplyr)
read.csv("WisconsinCancer.csv")
```


```{r}
fna.data <- "WisconsinCancer.csv"
wisc.df <-read.csv(fna.data, row.names=1)
View(wisc.df)

```
```{r}
wisc.data <- wisc.df[,-1]
```
Q1. How many observations are in this dataset?
```{r}
# Have a brief view first
View(wisc.df)
# How many row this data has?
nrow(wisc.df)
```
Q2. How many of the observations have a malignant diagnosis? 
```{r}
# How many people got diagnosis?
table(wisc.df$diagnosis)
# How many of them are malignant :(
sum(wisc.df$diagnosis=="M")
```
Q3. How many variables/features in the data are suffixed with _mean? 
```{r}
#1 way
length(grep("mean", colnames(wisc.df)))
#2 way
grepl("_mean", colnames(wisc.df)) %>% sum(.)
```

## Principal Component Analysis (PCA)
the main function in base R for PCA is 'prcomp()'. There is an important optional argument called 'scale'
```{r}
diagnosis <- wisc.df$diagnosis
wisc.data<- wisc.df[,-1]
```

```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
```

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale =TRUE)
# Look at summary of results
summary(wisc.pr) 

```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?


Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?


##Interpreting PCA results
```{r}
#plot biplot
biplot(wisc.pr)
```

```{r}
wisc.pr$rotation
```

```{r}
plot(wisc.pr$x[, 1:2], col = diagnosis, 
     xlab = "PC1", ylab = "PC2")
```


```{r}
plot(wisc.pr$x[, c(1,3)], col = diagnosis, 
     xlab = "PC1", ylab = "PC3")
```


```{r}
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis
library(ggplot2)
```

```{R}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]), method="ward.D2")
plot(wisc.pr.hclust)
```

```{r}

grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
```



```{R}
table(grps, diagnosis)
grps <- cutree(wisc.pr.hclust, k=2)
```



