---
title: "class 07"
author: "Phoebe LI"
date: "2/13/2022"
output: pdf_document
---

# Clusting with kmeans() and hclust()
we will begin by making up some data to cluster.

```{r}
tmp<-c(rnorm(30,3), rnorm(30.-3))
hist(tmp)
```


```{r}
x<-cbind(x=tmp,y=rev(tmp))
plot(x)
x
```

## Run Kmeans()
```{r}
K<- kmeans(x, centers = 2, nstart = 20)
K
```
Cluster membership
Q. what size is each cluster
```{r}
K$size
```
Q. cluster centers
```{r}
K$centers
```
>Q. membership vector

```{r}
K$cluster
```
 plot our data with the clusting result
 
```{r}
plot(x, col=K$cluster)
points(K$centers, col="blue", pch=16,cex=2)
```

## hclust()

Hierarchical Clustering
```{r}
hc<- hclust( dist(x) )
hc
```
 There is a cool and useful plot method for hcluster()
```{r}
plot(hc)
```
 
 
## 1. PCA of UK food data
 
Data import

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
x
```
 
>Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
## Complete the following code to find out how many rows and columns are in x?
dim(x)
nrow(x)
nrow(x)
```

```{r}
 ## Preview the first 6 rows
head(x,6)
```
```{r}
# Note how the minus indexing works

x <- x[,-1]
head(x)
dim(x)
```
```{r}
x <- read.csv(url, row.names=1)
head(x)
dim(x)
```
>Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

  I prefer the second one, becuase it makes it clear the function of the first row is name.
  
  
  Spotting major differences and trends
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```
>Q3: Changing what optional argument in the above barplot() function results in the following plot?

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

> Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?

The figure means the country you see the name on the vertical vs harizontal. 

```{r}
pairs(x, col=rainbow(10), pch=16)

```
> Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?

The data is more spread compare to others. They intake higher fresh potato than others. Other countries food consume are similar, but N.Ireland is more different than others.

## PCA to the rescue!
Do PCA of 17D food data. The main function in base R is called prcomp()
 
```{r}
#this function require tranpom of the data t(x)
pca<-prcomp(t(x))
summary(pca)
```
 
 
 The 'prcomp()' function return a list of object.
 
```{r}
plot(pca)
```
 The "PCA plot" aka a pca score plot is a plot of PC1 vs PC2. Basically using the new PCA axis to view our data.
 
```{r}
attributes(pca)
```
 
 
 
```{r}
pca$x
```
 
 
 >Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.
 
```{r}
plot(pca$x[,1], pca$x[,2],xlab="PC1", ylab="PC2")
text(pca$x[,1], pca$x[,2], labels = colnames(x))
```
>Q8. Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at start of this document.

```{r}
plot(pca$x[,1], pca$x[,2],xlab="PC1", ylab="PC2")
text(pca$x[,1], pca$x[,2], labels = colnames(x), col=rainbow(10))
```
 Below we can use the square of pca$sdev , which stands for “standard deviation”, to calculate how much variation in the original data each PC accounts for.
 
```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
## or the second row here...
z <- summary(pca)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```


## Digging deeper (variable loadings)
```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```
> Q9: Generate a similar ‘loadings plot’ for PC2. What two food groups feature prominantely and what does PC2 maninly tell us about?

It shows that fresh potato and soft drinks are two food groups feature prominantely.
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```
 
 
# Biplots
```{r}
## The inbuilt biplot() can be useful for small datasets 
biplot(pca)

```


## PCA of RNA-seq data
```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```
> Q10: How many genes and samples are in this data set?

```{r}
ncol(rna.data)
nrow(rna.data)
dim(rna.data)
```

```{r}
## Again we have to take the transpose of our data 
pca <- prcomp(t(rna.data), scale=TRUE)
 
## Simple un polished plot of pc1 and pc2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
```


This quick plot looks interesting with a nice separation of samples into two groups of 5 samples each. Before delving into the details of this grouping let’s first examine a summary of how much variation in the original data each PC accounts for:
```{r}
summary(pca)
```

A quick barplot summary of this Proportion of Variance for each PC can be obtained by calling the plot() function directly on our prcomp result object.
```{r}
plot(pca, main="Quick scree plot")
```
 Let’s make the above scree plot ourselves and in so doing explore the object returned from prcomp() a little further. We can use the square of pca$sdev, which stands for “standard deviation”, to calculate how much variation in the original data each PC accounts for:
```{r}
## Variance captured per PC 
pca.var <- pca$sdev^2

## Percent variance is often more informative to look at 
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```
 We can use this to generate our own scree-plot like this
```{r}
barplot(pca.var.per, main="Scree Plot", 
        names.arg = paste0("PC", 1:10),
        xlab="Principal Component", ylab="Percent Variation")
```
 Now lets make our main PCA plot a bit more attractive and useful…
 
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(rna.data)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"

plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))

text(pca$x[,1], pca$x[,2], labels = colnames(rna.data), pos=c(rep(4,5), rep(2,5)))
```
 
 
# Using ggplot
 
```{r}
library(ggplot2)

df <- as.data.frame(pca$x)

# Our first basic plot
ggplot(df) + 
  aes(PC1, PC2) + 
  geom_point()
```
If we want to add a condition specific color and perhaps sample label aesthetics for wild-type and knock-out samples we will need to have this information added to our data.frame:
```{r}
# Add a 'wt' and 'ko' "condition" column
df$samples <- colnames(rna.data) 
df$condition <- substr(colnames(rna.data),1,2)

p <- ggplot(df) + 
        aes(PC1, PC2, label=samples, col=condition) + 
        geom_label(show.legend = FALSE)
```
some polish
```{r}
p + labs(title="PCA of RNASeq Data",
       subtitle = "PC1 clealy seperates wild-type from knock-out samples",
       x=paste0("PC1 (", pca.var.per[1], "%)"),
       y=paste0("PC2 (", pca.var.per[2], "%)"),
caption="BIMM143 example data") +
     theme_bw()
```

## Optional: Gene loadings
```{r}
loading_scores <- pca$rotation[,1]

## Find the top 10 measurements (genes) that contribute
## most to PC1 in either direction (+ or -)
gene_scores <- abs(loading_scores) 
gene_score_ranked <- sort(gene_scores, decreasing=TRUE)

## show the names of the top 10 genes
top_10_genes <- names(gene_score_ranked[1:10])
top_10_genes 

```

