---
title: "class12"
author: "Phoebe LI"
date: "2/25/2022"
output: pdf_document
---
1. input our counts and metadata files.
-check the format and fix if necessary

2. Run differential expression analysis
-setup that object required by deseq()
-run deseq

3. Add annotation
-gene names and something else 

4. Volcano plot

5. Pathway analysis

6. Save results

7. Go have fun!

```{r setup, message=FALSE}
library(DESeq2)
library(ggplot2)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(gage)
library(gageData)
library(pathview)
```

#input counts and metadata
```{r readcvs}
countdf<- read.csv("GSE37704_featurecounts.csv", row.names = 1)
colData<- read.csv("GSE37704_metadata.csv")

#head(countdf)
countData<- countdf[,-1]
head(countData)

```

```{r}
#all(colData$id==colnames(countData))
```

```{r}
#head(countData)
counts <- countData[rowSums(countData) != 0, ]
head(counts)
#head(countData)
```

## Principle component Analysis

```{r}
pca<-prcomp(t(counts), scale=TRUE)
summary(pca)
```
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
```

```{r}
#plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", col=colData$condition)
plot(pca$x[,1:2],col=as.factor(colData$condition),pch=16)
```



## DESeq Analysis

```{r}
dds<- DESeqDataSetFromMatrix(countData = counts,
                       colData = colData,
                       design = ~condition)
```
Run our DESeq analysis
```{r}
dds<-DESeq(dds)
res<-results(dds)
head(res)
```


## Add Annotation

```{r}
columns(org.Hs.eg.db)

res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(counts), 
                    keytype="ENSEMBL",
                    column="SYMBOL",
                    multiVals="first")

res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(counts),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

res$name =   mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="GENENAME",
                    multiVals="first")
head(res, 10)
```


```{r}
 plot( res$log2FoldChange, -log(res$padj) )
```


```{r}
# Make a color vector for all genes
mycols <- rep("gray", nrow(res) )

# Color red the genes with absolute fold change above 2
mycols[ abs(res$log2FoldChange) > 2 ] <- "red"

# Color blue those with adjusted p-value less than 0.01
#  and absolute fold change more than 2
inds <- ((res$pvalue)<0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

plot( res$log2FoldChange, -log(res$padj), col=mycols, xlab="Log2(FoldChange)", ylab="-Log(P-value)" )
```
## pathway analysis

```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```


```{r}
library(pathview)
library(gage)
library(gageData)
data(sigmet.idx.hs)
data(kegg.sets.hs)

kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]
head(kegg.sets.hs, 3)
```


```{r}
keggres = gage(foldchanges, gsets=kegg.sets.hs)
attributes(keggres)
```

```{r}
head(keggres$less)
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```

## Section 3. Gene Ontology (GO)

```{r}
data (go.sets.hs)
data (go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```




```{r}
res = res[order(res$pvalue),]
write.csv(res, file="deseq_results.csv")
```

