---
title: "class6"
author: "Phoebe LI"
date: "2/13/2022"
output: html_document
---

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

We can write text of course just like any file. We
can **style text
```{r}

```

Do :

- this
-and that
-and another thing

This is more text
and this a new line

---

we can include some code:

```{r}
plot(1:10)
```

```{r}
# This is a comment and will not be to R
plot(cars)
```
```{r}

```

## Time to write a function

> **Q1**. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://urldefense.proofpoint.com/v2/url?u=https-3A__tinyurl.com_gradeinput-25C3-25A2&d=DwIGAw&c=-35OiAkTchMrZOngvJPOeA&r=1eeYiVnpi6GuTHHj4Q58Ig&m=DRKZ8KtZXkjKttHZO3O7EqYeh2fRyq8XmhIhuLo69AZN-Zpxe8jeU2QLHAao2q4S&s=wtn9GEwNwNfCQtVcCHceYdY5yd9FyU31iesjgte30vo&e= ?? [3pts]






First i want to find the lowest score. I can use the **min()** to find it and the **which.min()** function to find where it is (i,e its position in the vector).

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

```
```{r}
which.min(student1)
```
```{r}
student1[ -which.min(student1)]
```
i can use minus to get everything in the vector but the lowest score. 
 
Now i can call for thr **mean()** function to get the average.
```{r}
mean(student1[ -which.min(student1)])
```
```{r}
student2
which.mean(student2)])
```

```{r}
mean(student2, na.rm = TRUE)
```
```{r}
student3
```

one great idea is to change or replace the NA values with zero. lets do it.

The internet gods told me to try this 
```{r}
which(is.na(student2))
```
The is.na() function returns a logical vector where TRUE elements indicate the presence of NA values.
```{r}
is.na(student2)
```


```{r}
student2[ is.na]
```

lets replace NAs with zero

```{r}
student.prime <- student3
student.prime[ is.na(student.prime)] = 0
stuent.prime
```

ok we are so close lets put these bits

```{r}
student.prime <- student2
student.prime[ is.na(student.prime)] = 0
mean(student.prime[ -which.min(student.prime)])
```
```{r}
x <- 1:5
x
x[2] = 100
x
x[3] = 200
```
hOW about student3?
```{r}
student.prime <- student3
student.prime[ is.na(student.prime)] = 0
mean(student.prime[ -which.min(student.prime)])
```

great we got it. this works. lets simplify this and make it clear


we can make the object names more clear

```{r}
x <- student3
x[ is.na(student.prime)] = 0
mean(x[ -which.min(student.prime)])
```

Doh! eeijt barry entered the wrong code
```{r}
student4 <- c(100, NA, 90, "90", 90, 90, 97, 80)
```

```{r}
as.numeric(student4)
```
```{r}
as.numeric(student4)
```

```{r}
x <- student4
x <- as.numeric(x)
x[ is.na(x)] = 0
mean(x[ -which.min(x)])
```

noe finally we canwite our function:
Al functions have at least 3 things 
a name, imout args abd a body.

```{r}
grade <- function(x) {
  x <- as.numeric(x)
x[ is.na(x)] = 0
mean(x[ -which.min(x)])
}
```

```{r}
grade(student1)
```

First we got to read the gradebook for teh class.

```{r}
gradebook <- "https://urldefense.proofpoint.com/v2/url?u=https-3A__tinyurl.com_gradeinput&d=DwIGAw&c=-35OiAkTchMrZOngvJPOeA&r=1eeYiVnpi6GuTHHj4Q58Ig&m=DRKZ8KtZXkjKttHZO3O7EqYeh2fRyq8XmhIhuLo69AZN-Zpxe8jeU2QLHAao2q4S&s=rauPa6Wl4Ax1RnpvkygjFsIjgO_qG0KkCmqz8_hV56I&e= "
scores <- read.csv(gradebook, row.names = 1)
scores
```

We are goign to use the supper useful **apply()** function to grade all the students with our **grade()** function.

```{r}
apply(scores, 1, grade)
```
```{r}
apply(scores, 1, grade)
```

```{r}
which.max(ans)
```


question 3: which question was the toughest
here we will 

```{r}
apply(scores,2,mean, na.rm=TRUE)
```

Replace or mask NA values to zero
```{r}
mask <- scores
```

```{r}
mask <- scores
mask[is.na(mask)] = 0
mask
```

```{r}
apply(mask,2,mean)
```

```{r}
cor(mask$hw5, ans)
```

I can call the **cor()** for every homework and get a value for each but that sucks. lets use **()** and do them all in one go.

```{r}
apply(mask,2,cor, ans)
```

