---
title: "class06BGGN213"
author: "Phoebe LI"
date: "2/7/2022"
output: pdf_document
---

Here we will write a function to grade some student homework
We will start with a more simple example
```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90) 
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80) 
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

```


```{r}
student1
```
The regular average will be returned by mean function.
```{r}
mean(student1)
```
To find the position of the smallest value

```{r}
which.min(student1)
```
so this will give us the min value
```{r}
# The first solution
mean(student1[-which.min(student1)])

```
How about NA? it is a bad idea to use na.rm=T
```{r}
# For students who have NA as their score
# Change NA=0 
x<- student2
x

x[is.na(x)]<-0
x
mean(x)
```

Find the average for student 3
```{r}
x<- student3
x[is.na(x)]<-0
mean(x[-which.min(x)])

```
## Let's make a function!! REMEBER TO EXTACT THE FUNCTION

```{r}
grade <- function(x) {
    x[is.na(x)]<-0
    mean(x[-which.min(x)])
}
```
Let's grade students 1,2,3
```{r}
grade(student1)
grade(student2)
grade(student3)
```
## We can grade the Whole class now
> Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. 

```{r}
#' Map missing HW, NA value to 0
#' We exclude the lowest one 
#'
#' @param x Numeric vector of home
#' @return Average score
#' @example
#'   student<-c(100, NA, 90, 80)
#'   grade(student)
#' 
grade <- function(x) {
    #map missing HW, NA value to 0
    x[is.na(x)]<-0
    # We exclude the lowest one   
    mean(x[-which.min(x)])
}
```
>Your final function should be adquately explained with **code comments** and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
url <- "https://tinyurl.com/gradeinput"
gradebook<- read.csv(url, row.names=1)
```

```{r}
apply(gradebook, 1, grade)
```

> Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook? [3pts]

```{r}
results <- apply(gradebook, 1, grade)
sort(results, decreasing = TRUE)
```
Or you can find the one has highest score
```{r}
which.max(results)
```

>Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]

```{r}
# Let's try average
hw.ave<-apply(gradebook, 2, mean, na.rm=TRUE)
which.min(hw.ave)
```

```{r}
hw.med<-apply(gradebook, 2, median, na.rm=TRUE)
which.min(hw.med)
```
So, which one I should use??? Let's plot it!!!

```{r}
boxplot(gradebook)
```



>Q4. Optional Extension: From your analysis of the gradebook, which homework was most predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

```{r}
gradebook[ is.na(gradebook)] <-0
cor(results, gradebook$hw2)
```
```{r}
apply(gradebook, 2, cor, x=results)
```

>Q5. Make sure you save your Rmarkdown document and can click the “Knit” button to generate a PDF foramt report without errors. Finally, submit your PDF to gradescope. [1pt]

